---
title: "Comparing Data"
author: "Chi Kit Yeung"
date: "2022-08-05"
output: html_document
---
## Operators

---

### Custom 'NOT IN' Operator

R has a built in `%in%` operator that's useful for comparing values that's similar to SQL's `LIKE` operator.  However, it doesn't have a built in `NOT LIKE` operator like SQL does which is useful in some cases.

```{r}
# Defining the operator

`%!in%` <- Negate(`%in%`)
```

Example:

```{r}
fav_fruits <- c('apple', 'oranges', 'bananas', 'grape')
shopping_list <- c('bananas', 'persimmon', 'peach', 'apple', 'custard apple')

# Normal `%in%` operator use
fav_fruits[fav_fruits %in% shopping_list]
```
Above we can see two of our favorite fruits being on the shopping list

Next, using the custom defined `%!in%` operator
```{r}
# Favorite fruits not being bought

fav_fruits[fav_fruits %!in% shopping_list]
```
```{r}
# Not so favorite fruits being bought :(

shopping_list[shopping_list %!in% fav_fruits]
```

---

### Intersect and SetDiff

```{r}
intersect(fav_fruits, shopping_list)
```


```{r}
setdiff(fav_fruits, shopping_list)
```


## Uncategorized

### Regex

Regular expressions can be utilized using the str_detect() function.

```{r}
library(tidyverse)

str_detect(fav_fruits, "a..le")
```
```{r} 
# Matching
fav_fruits[str_detect(fav_fruits, "a..le")]
```

```{r}
# Not matching
fav_fruits[!str_detect(fav_fruits, "a..le")]
```

### Unique

Getting unique values from a list using the `unique()` function.

```{r}
# A lot of dupes here, I just want the unique values!
locales <- c("ar_SA", "ar_SA", "de_DE", "de_DE", "de_DE", "en_AU", "es_ES", "es_ES", "es_ES", "es_ES", "es_MX", "es_MX", "es_MX", "es_MX", "es_US", "es_US", "es_US", "es_US", "es_US", "es_US", "fr_FR", "he_IL", "he_IL", "it_IT", "it_IT", "it_IT", "nb_NO", "nb_NO", "ru_RU", "ru_RU", "ru_RU", "sv_SE", "sv_SE", "sv_SE", "tr_TR", "tr_TR", "tr_TR")

locales
```

```{r}
unique(locales)
```

