---
title: "Importing Data"
author: "Chi Kit Yeung"
date: "2022-08-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Importing Multiple CSVs from a Directory

A lot of times I find myself with data sources that are stored in multiple separate csv files, this method seems to be the best one I've found so far to import them all into a single dataframe.

```{r eval=FALSE}
# Importing the data
# First define the path where the data is stored
path <- "../data/"
filenames <- list.files(path = path)

## Below is to import a specific range of files in the directory
# files <- list.files(path = path)
# filenames <- files[c((length(files)-7):length(files))]


## Into a list
all_df <- lapply(filenames, function(i) {
  i <- paste(path,i,sep="")
  read.csv(i, header=FALSE)
  })
filenames <- gsub("-","_",filenames)
names(all_df) <- gsub(".csv","",filenames)

## Joining all the data frames
joined_df <- all_df %>% 
  reduce(full_join)

## Reduce function above appends a new placeholder column name and pushes the actual name to first row
names(joined_df) <- joined_df[1,]
df <- joined_df %>% 
  slice(-1)
```

The disadvantage of this method is that all data types will be converted to `char` so an additional step will be needed to convert your data into the right types.


